### 2023-02-14 intro.qmd
update

![Illustration of the global carbon cycle (numbers represent a global average for the decade 2012-2021). Arrows and circles depict, respectively, carbon fluxes in Pg C year$^{-1}$ and carbon content in Pg C. Figure taken from @Friedlingstein2022-XX.](image/intro/global_carbon_cycle.png){#fig-global_carbon_cycle}

@@

Thankfully, more than half of this anthropogenic $CO_{2}$ has been absorbed by terrestrial ecosystems and by the ocean [@Friedlingstein2022-XX], forming major carbon sinks.

@@

The ocean is known to play a key role in the fast carbon cycle due to its size and rapid exchange of $CO_{2}$ with the atmosphere, controlling atmospheric $CO_{2}$ levels on timescales of decades to millennia [@Sarmiento2006-ec; @DeVries2022-of]. Currently, the ocean absorbs \~2.5 Pg C year$^{-1}$, which represents about 25% of annual anthropogenic $CO_{2}$ emissions [@Canadell2021-ds; @Friedlingstein2022-XX]. In order to understand how the ocean absorbs the latter, one must take a look at seawater carbonate chemistry.

@@

Also known as the soft-tissue pump [@Volk1985-oc], the biological carbon pump (BCP), is defined by @Ipcc2019-qc as "a series of ocean processes through which inorganic carbon is fixed as organic matter by photosynthesis (@eq-photosynthesis) in sunlit surface water and then transported to the ocean interior, and possibly the sediment, resulting in the storage of carbon".

@@

This sequestration (storage) of carbon is usually defined as the long-term removal of $CO_{2}$ (or other forms of carbon) from the atmosphere on climatically significant time scales [e.g. @Boyd2019-ez; @Ipcc2019-qc]. While this definition seems logical, it does not specify precisely what is meant by "climatically significant time scales". In what follows, I will use the 100-year time horizon to refer to carbon sequestration (unless specified otherwise) as used by other BCP studies [e.g. @Lampitt2008-kq; @Passow2012-zh].

The amount of carbon that is sequestered by the BCP mainly depends its strength and its efficiency. In this manuscript, the strength of the BCP is defined as "the magnitude of the particulate organic carbon (POC) flux at a reference export depth" [@Claustre2021-tx] and its efficiency (also known as transfer efficiency) is defined as the fraction of POC flux transported to some reference depth below the export depth [@Buesseler2020-sn; @Claustre2021-tx]. The latter is often set at 100 m but @Buesseler2020-sn strongly emphasized that carbon flux-related studies should use the base of the euphotic zone (EZ, where there is enough light for primary production) as the export depth to clearly differentiate the layer where POC is produced (i.e. via photosynthesis) and where POC is remineralized (or respired), that is organic matter converted back to inorganic forms (i.e. nutrients and DIC) by heterotrophic organisms [@Siegel2022-wd].

In comparison, the recent work of @Nowicki2022-no suggests that the BCP exports on average 10.2 Pg C year$^{-1}$, sequesters \~1,300 Pg C with a mean sequestration time of 127 years and buries 0.4 Pg C year$^{-1}$ in the sediment that will remain sequestered for millions of years. Therefore, when talking about natural-based solutions for carbon sequestration, one should keep in mind that even though the BCP does not contribute to the absorption of anthropogenic $CO_{2}$ *per se*, it remains the main process that maintains the DIC gradient.

@@

The BCP is a natural-based solution for carbon sequestration by definition (@eq-photosynthesis). However, natural-based solutions for carbon sequestration are generally aligned with the concept of 'blue carbon' defined by @Ipcc2019-qc as "all biologically-driven carbon fluxes and storage in marine systems that are amenable to management" which usually encompasses only coastal ecosystems [@Ipcc2019-qc; @Lovelock2019-dr]. Coastal blue carbon ecosystems (e.g. mangroves, tidal marshes, seagrass meadows) are thought to store in their living biomass (years to decades) and sediment (millenia) more than 30 Pg C with an estimated absorption of \~0.3 Pg C year$^{-1}$ [@Macreadie2021-rb] though those numbers should be taken with caution because there are major uncertainties in the distribution of seagrass meadows and tidal marshes [@Macreadie2021-rb] as well as in carbon burial rates [@Williamson2022-ud]. In spite of that, the current expectation of additional carbon removal by coastal blue carbon by 2100 should be around 25-30 Pg C [@Gattuso2021-cr; @Williamson2022-ud], that is the equivalent of 2.5-3 years of current anthropogenic emissions [@Friedlingstein2022-XX].
:::

According to @Sarmiento2006-ec, the BCP would thus be responsible for the remaining 70% of the observed DIC gradient. However, it should be emphasized that in contrast to the solubility pump, the BCP is not responsible for the absorption of excess atmospheric $CO_{2}$, released by anthropogenic activities. First, as explained in @Sarmiento2006-ec, atmospheric $CO_{2}$ levels were stable for several thousand years before the industrial revolution started, therefore it is assumed that the BCP was operating at steady state. Second, photosynthesis is mostly controlled by light and nutrients availability, not by $CO_{2}$.

@@

@@

Regarding the latter, @Boyd2019-ez published a paper in which they review the BCP by adding new biological and physical mechanisms[^intro-2] that inject suspended and sinking particles to depth (@fig-global_carbon_pumps). A description of those mechanisms is provided in the next subsections.

@@

### Gravitational pump {#sec-BGP}

The most well known pump is the gravitational carbon pump (BGP). It consists in the gravitational settling (i.e. sinking) of POC[^intro-3] to the deep ocean and it is the main contributor (60-70%) of the export flux [@Boyd2019-ez; @Nowicki2022-no]. Historically, the downward POC flux was directly measured using sediment traps [@Martin1987-um; @Honjo2008-kn] and indirectly measured using thorium-234 ($^{234}Th$) deficits as a POC flux proxy [e.g. @Buesseler2006-og]. More recently, the advent of underwater cameras such as the Underwater Vision Profiler 5 (UVP5) [@Picheral2010-to] allowed the direct measurement of particle size spectra (PSS) that could then be converted to carbon fluxes [e.g. @Guidi2008-vp].

@@

The BGP was long thought to be the only component of the BCP therefore new instruments are being tested to decouple the downward POC flux signal from the contribution of other pumps (@sec-BGC-sensors).

### Mixing pump {#sec-MixingPump}

The mixing pump is a physically mediated particle injection pump (PIP). It consists of three separate pumps: the mixed layer pump, the eddy subduction pump and the large-scale subduction pump [@Boyd2019-ez]. Contrarily to the BGP, all three pumps also transport DOC to depth where it is injected along with suspended (i.e. neutrally buoyant) and slow-sinking POC below the EZ on different spatial scales and timescales.

The mixed layer pump can be defined as a process through which carbon is exported below the EZ due to variation in the depth of the mixed layer [@DallOlmo2016-ex]. The eddy subduction pump is associated with fronts and eddies that create a strong vertical circulation that transport particles on timescales of days across horizontal spatial scales of 1-10 km [@Omand2015-mf; @Boyd2019-ez]. The large-scale (100-1000 km) subduction pump corresponds to the slow advective transport of carbon associated with the meridional overturning circulation and Ekman pumping [@Levy2013-ro; @Claustre2021-tx; @Siegel2022-wd]. The last two pumps are both 2D spatial processes (i.e. they transport POC and DOC vertically and laterally).

The mixing pump would represent at least 20% of the export flux according to @Nowicki2022-no although large uncertainties remain (see Figure 2C in the paper of @Boyd2019-ez).

### Migrant pump {#sec-MigrantPump}

The migrant pump, also known as the vertical migration pump (VMP), is a biologically mediated PIP [@Boyd2019-ez]. It consists on the vertical migration of zooplankton or fish [@Steinberg2017-ri; @Saba2021-ew] that directly inject POC (e.g. fecal pellets) below the EZ at the depth of their migration. This pump can act on daily and seasonal timescales. In the case of the diel (i.e. daily) migration pump, animals feed at the surface at night and then migrate to depth at sunrise to avoid predation [@Siegel2022-wd], metabolizing injected food at depth[^intro-4], by-passing the EZ directly [@Steinberg2017-ri]. In the case of the seasonal migration pump, carbon-rich lipids accumulated in the EZ during summer are catabolized by hibernating copepods at depth, between 600 m and 1,400 m, where they shunt carbon below the permanent pycnocline [@Jonasdottir2015-ob]. This pump is also referred to as the seasonal lipid pump [@Jonasdottir2015-ob].

>>>>

![Illustration of the global carbon cycle (numbers represent a global average for the decade 2012-2021). Arrows and circles depict, respectively, carbon fluxes in Pg C year$^{-1}$ and carbon content in Pg C. Figure taken from @Friedlingstein2022-th.](image/intro/global_carbon_cycle.png){#fig-global_carbon_cycle}

@@

Thankfully, more than half of this anthropogenic $CO_{2}$ has been absorbed by terrestrial ecosystems and by the ocean [@Friedlingstein2022-th], forming major carbon sinks.

@@

The ocean is known to play a key role in the fast carbon cycle due to its size and rapid exchange of $CO_{2}$ with the atmosphere, controlling atmospheric $CO_{2}$ levels on timescales of decades to millennia [@Sarmiento2006-ec; @DeVries2022-of]. Currently, the ocean absorbs \~2.5 Pg C year$^{-1}$, which represents about 25% of annual anthropogenic $CO_{2}$ emissions [@Canadell2021-ds; @Friedlingstein2022-th]. In order to understand how the ocean absorbs the latter, one must take a look at seawater carbonate chemistry.

@@

Also known as the soft-tissue pump [@Volk1985-oc], the biological carbon pump (BCP), is defined by @Ipcc2019-qc as "a series of ocean processes through which inorganic carbon is fixed as organic matter by photosynthesis (@eq-photosynthesis) in sunlit surface water and then transported to the ocean interior (i.e. exported flux), and possibly the sediment, resulting in the storage of carbon".

@@

In order to estimate the carbon export flux, one needs to set a depth horizon [@Siegel2022-wd]. Historically, the latter has often been set at 100 m but @Buesseler2020-sn strongly emphasized that carbon flux-related studies should now use the base of the euphotic zone (EZ, where there is enough light for primary production) as the export depth to clearly differentiate the layer where particulate organic carbon (POC) is produced (i.e. via photosynthesis) and where POC is consumed (i.e. remineralized). This remineralization (or respiration) consists in the conversion of organic matter back to inorganic forms (i.e. nutrients and DIC) by heterotrophic organisms [@Siegel2022-wd].

According to @Sarmiento2006-ec, the BCP would thus be responsible for the remaining 70% of the observed DIC gradient. Without it, there would be no production of organic matter in the EZ (consuming DIC) and no remineralization in the ocean interior (producing DIC) resulting in a weaker DIC gradient. In fact, @DeVries2022-of calculated that such a decrease in the DIC gradient would reduce the ocean carbon inventory by roughly 10% and increase atmospheric $CO_{2}$ levels by a factor 2. Similarly, @Kwon2009-ir showed that a deepening of the remineralization depth (i.e. where 63% of sinking carbon is respired) would decrease atmospheric $CO_{2}$ concentrations due to the redistribution of respired carbon from intermediate waters to bottom waters, enhancing the DIC gradient.

The sinking carbon that is not remineralized is said to be sequestered. The sequestration (storage) of carbon is usually defined as the long-term removal of $CO_{2}$ (or other forms of carbon) from the atmosphere on climatically significant time scales [@Boyd2019-ez; @Ipcc2019-qc; @Siegel2022-wd]. While this definition seems logical, it does not specify precisely what is meant by "climatically significant time scales". In what follows, I will use the 100-year time horizon to refer to carbon sequestration (unless specified otherwise) as used by other BCP studies [e.g. @Lampitt2008-kq; @Passow2012-zh].

The amount of carbon that is sequestered by the BCP mainly depends on its strength and its efficiency. In this manuscript, the strength of the BCP is defined as "the magnitude of the POC flux at a reference export depth" [@Claustre2021-tx] and its efficiency (also known as transfer efficiency) is defined as the fraction of POC flux transported to some reference depth below the export depth [@Buesseler2020-sn; @Claustre2021-tx].

@@

The BCP is a natural-based solution for carbon sequestration by definition (@eq-photosynthesis). However, natural-based solutions for carbon sequestration are generally aligned with the concept of 'blue carbon' defined by @Ipcc2019-qc as "all biologically-driven carbon fluxes and storage in marine systems that are amenable to management" which usually encompasses only coastal ecosystems [@Ipcc2019-qc; @Lovelock2019-dr]. Coastal blue carbon ecosystems (e.g. mangroves, tidal marshes, seagrass meadows) are thought to store in their living biomass (years to decades) and sediment (millenia) more than 30 Pg C with an estimated absorption of \~0.3 Pg C year$^{-1}$ [@Macreadie2021-rb] though those numbers should be taken with caution because there are major uncertainties in the distribution of seagrass meadows and tidal marshes [@Macreadie2021-rb] as well as in carbon burial rates [@Williamson2022-ud]. In spite of that, the current expectation of additional carbon removal by coastal blue carbon by 2100 should be around 25-30 Pg C [@Gattuso2021-cr; @Williamson2022-ud], that is the equivalent of 2.5-3 years of current anthropogenic emissions [@Friedlingstein2022-th].
:::

Estimates of global carbon export range between 5 and 12 Pg C year$^{-1}$ [@Dunne2007-mi; @Henson2011-mt; @Siegel2014-wg; @DeVries2017-wu; @Nowicki2022-no] with only 0.02-0.2 Pg C year$^{-1}$ reaching the sediment [@Dunne2007-mi; @Hayes2021-gg] where carbon will be sequestered for millions of years. Almost all of the exported organic carbon will therefore be remineralized and sequestered away from the atmosphere on time scales ranging from months to millennia [@Siegel2022-wd].

Nonetheless, it should be emphasized that in contrast to the solubility pump, the BCP is not responsible for the absorption of excess atmospheric $CO_{2}$, released by anthropogenic activities. First, as explained in @Sarmiento2006-ec, atmospheric $CO_{2}$ levels were stable for several thousand years before the industrial revolution started, therefore it is assumed that the BCP was operating at steady state. Second, photosynthesis is mostly controlled by light and nutrients availability, not by $CO_{2}$.

@@

<!-- he recent work of @Nowicki2022-no suggests that the BCP exports on average 10.2 Pg C year$^{-1}$, sequesters ~1,300 Pg C with a mean sequestration time of 127 years and buries 0.4 Pg C year$^{-1}$ in the sediment that will remain sequestered for millions of years. Therefore, when talking about natural-based solutions for carbon sequestration, one should keep in mind that even though the BCP does not contribute to the absorption of anthropogenic $CO_{2}$ per se, it remains the main process that maintains the DIC gradient -->

@@

Regarding the latter, @Boyd2019-ez published a paper in which they review the BCP by adding new biological and physical mechanisms[^intro-2] that inject suspended and sinking particles to depth (@fig-global_carbon_pumps). A better understanding of those mechanisms (see next subsections) could help to close the carbon budget.

@@

### Biological gravitational pump  {#sec-BGP}

The most well known pump is the gravitational carbon pump (BGP). It consists in the gravitational settling (i.e. sinking) of POC[^intro-3] to the deep ocean and it is the main contributor (60-70%) of the export flux [@Boyd2019-ez; @Nowicki2022-no]. Historically, the downward POC flux was directly measured using sediment traps [@Martin1987-um; @Honjo2008-kn] and indirectly measured using thorium-234 ($^{234}Th$) deficits as a POC flux proxy [e.g. @Buesseler2006-og]. More recently, the advent of underwater cameras such as the Underwater Vision Profiler 5 (UVP5) [@Picheral2010-to] allowed the direct measurement of particle size distributions (PSD) that could then be converted to carbon fluxes [e.g. @Guidi2008-vp].

@@

### Physically mediated particle injection pump {#sec-MixingPump}

The physically mediated particle injection pump consists of three separate pumps: the mixed layer pump, the eddy subduction pump and the large-scale subduction pump [@Boyd2019-ez]. Contrarily to the BGP, all three pumps also transport DOC to depth where it is injected along with suspended (i.e. neutrally buoyant) and slow-sinking POC below the EZ on different spatial scales and timescales.

The mixed layer pump can be defined as a 1D process through which organic carbon is exported below the EZ when there is a deepening of the mixed layer followed by the formation of a new shallower mixed layer, isolating the exported carbon from the surface [@DallOlmo2016-ex]. This pump can occur at different time scales but is mainly related to the building of the summer stratification in mid and high latitude regions where the seasonal variations of the mixed layer are the strongest [@DallOlmo2016-ex; @Boyd2019-ez].

The eddy subduction pump is associated with fronts and eddies that create a strong vertical circulation that transport particles on timescales of days across horizontal spatial scales of 1-10 km \[@Omand2015-mf; @Resplandy2019-nx\]. The large-scale (100-1000 km) subduction pump corresponds to the slow advective transport of carbon associated with the meridional overturning circulation and Ekman pumping \[@Levy2013-ro; @Claustre2021-tx; @Siegel2022-wd\]. The last two pumps are both 2D spatial processes (i.e. they transport POC and DOC vertically and laterally).

The physically mediated particle injection pump would represent at least 20% of the export flux according to @Nowicki2022-no although large uncertainties still remain (see Figure 2C in the paper of @Boyd2019-ez).

### Biologically mediated particle injection pump {#sec-MigrantPump}

The biologically mediated particle injection pump is usually referred to as the vertical migration pump (VMP) [@Boyd2019-ez]. It consists on the vertical migration of zooplankton or fish [@Steinberg2017-ri; @Saba2021-ew] that directly inject POC (e.g. fecal pellets) below the EZ at the depth of their migration. This pump can act on daily as well as on seasonal timescales. In the case of the diel (i.e. daily) migration pump, animals feed at the surface at night and then migrate to depth during daytime to avoid visual predation [@Siegel2022-wd], metabolizing injected food at depth[^intro-4], by-passing the EZ directly [@Steinberg2017-ri]. In the case of the seasonal migration pump, carbon-rich lipids accumulated in the EZ during summer are catabolized by hibernating copepods at depth, between 600 m and 1,400 m, where they shunt carbon below the permanent pycnocline [@Jonasdottir2015-ob]. This pump is also referred to as the seasonal lipid pump [@Jonasdottir2015-ob].
